Mixing time estimation in reversible Markov chains from a single sample path

### 2015-06-02 abstract.tex
abstract

%!TEX root =  matrix-est.tex
This article provides the first procedure for computing a fully
data-dependent interval that traps the mixing time $\tmix$ of a finite
reversible ergodic Markov chain at a prescribed confidence level.
The interval is computed from a single finite-length sample
path from the Markov chain, and does not require the 
knowledge of any parameters of the chain.
%bounds on the mixing time itself.
This stands in contrast to previous approaches, which either 
only provide point estimates, or require a
reset mechanism, or additional prior knowledge.
%Most notably, 
%the procedure produces a fully empirical confidence
%interval (i.e., it depends only on the sample path and a given
%confidence parameter $\delta$) that is always valid with probability
%$1-\delta$; and 
The width of the interval converges to zero at a
$\sqrt{n}$ rate, where $n$ is the length of the sample path.
Upper and lower bounds are given on the number of samples required to
achieve constant-factor multiplicative accuracy.
\todoc{We may need to add some explanation to connect $\gap$ estimates
to $\tmix$ estimates.}
The lower bounds indicate that, unless further restrictions are placed
on the chain, no procedure can achieve this accuracy level before
seeing each state at least $\Omega(\tmix)$ times.
Future directions of research are identified.
\todoc{Such as: closing the gap. Extension to large state spaces, 
based on additional assumptions. Even if $\gap$ is large, we need many samples.
Maybe we need even stronger assumptions, like the decay rate of eigenvalues.
}
%If $\gap$ denotes the absolute spectral gap of the unknown Markov chain
%and the chain's stationary distribution is minorized by $\pimin$,
%our procedure requires $\tilde{O}( \frac{1}{\pimin\gap^3})$ observations
% to estimate the mixing time up to a constant multiplicative accuracy, 
%while we also establish that any algorithm to achieve the same
%goal would need $\Omega( d\log(d)/\gap + 1/\pimin)$ samples.

>>>>

%!TEX root =  matrix-est.tex
This article provides the first procedure for computing a fully
data-dependent interval that traps the mixing time $\tmix$ of a finite
reversible ergodic Markov chain at a prescribed confidence level.
The interval is computed from a single finite-length sample
path from the Markov chain, and does not require the 
knowledge of any parameters of the chain.
%bounds on the mixing time itself.
This stands in contrast to previous approaches, which either 
only provide point estimates, or require a
reset mechanism, or additional prior knowledge.
%Most notably, 
%the procedure produces a fully empirical confidence
%interval (i.e., it depends only on the sample path and a given
%confidence parameter $\delta$) that is always valid with probability
%$1-\delta$; and 
The width of the interval converges to zero at a
$\sqrt{n}$ rate, where $n$ is the length of the sample path.
Upper and lower bounds are given on the number of samples required to
achieve constant-factor multiplicative accuracy.
\todoc{We may need to add some explanation to connect $\gap$ estimates
to $\tmix$ estimates.}
The lower bounds indicate that, unless further restrictions are placed
on the chain, no procedure can achieve this accuracy level before
seeing each state at least $\Omega(\tmix)$ times on the average.
Future directions of research are identified.
\todoc{Such as: closing the gap. Extension to large state spaces, 
based on additional assumptions. Even if $\gap$ is large, we need many samples.
Maybe we need even stronger assumptions, like the decay rate of eigenvalues.
}
%If $\gap$ denotes the absolute spectral gap of the unknown Markov chain
%and the chain's stationary distribution is minorized by $\pimin$,
%our procedure requires $\tilde{O}( \frac{1}{\pimin\gap^3})$ observations
% to estimate the mixing time up to a constant multiplicative accuracy, 
%while we also establish that any algorithm to achieve the same
%goal would need $\Omega( d\log(d)/\gap + 1/\pimin)$ samples.
